---
title: "McKinsey Hackathon"
author: "Ida Davidsson"
date: "20/01/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Loading the data and the required libraries
```{r data and library setup}
conversions = read.csv("train.csv", stringsAsFactors=FALSE)
test_df = read.csv("test.csv")

library(tidyverse)
library(eeptools)
library(lubridate)
library(caret)
library(MASS)
library(randomForest)
```


Time to do some Feature Engineering.. I start with converting the Date of Birth (DOB) to age in years, as this makes binning easier (15 entires have NA age, so I set these to 0 as they are such a small amount and this will make the binning process easier). I then bin age and salary, and create a new variable called debt_to_income, which is the maonthly income post EMI payments divided by the loan amount.
```{r feature engineering}
# converting DOB to age in years
conversions = conversions %>%
  mutate(DOB = as.Date(conversions$DOB, "%d/%m/%y"))

conversions$year <- format(as.Date(conversions$DOB), "%Y")
conversions$year = as.numeric(conversions$year)

date = conversions$DOB

conversions$age = 2018 - year(date)
conversions$age = abs(conversions$age)

conversions$age[is.na(conversions$age)] = 0

# binning age into 3 categories (based on "middle age" being considered 45-65)
conversions = conversions %>%
  mutate(age_bracket = ifelse(age >= 45 & age <= 65, "middle_aged",
                         ifelse(age >= 18 & age < 45, "young_adult", "senior"))
  )

# binning income into 3 categories ($22,865.10 being the US federal govt for low income, and $100,000 being considered high income)
conversions = conversions %>%
  mutate(income_level = ifelse(Monthly_Income <= 1905, "low", 
                          ifelse(Monthly_Income > 1905 & Monthly_Income < 8334, "middle", "high")))

# Debt to income (after withdrawing current EMI)
conversions = conversions %>%
  mutate(income_post_EMI = Monthly_Income - Existing_EMI)

conversions = conversions %>%
  mutate(debt_to_income = conversions$income_post_EMI/conversions$Loan_Amount)

conversions$debt_to_income[is.na(conversions$debt_to_income)] = 0

# Binning of debt_to_income
conversions = conversions %>%
  mutate(debt_level = ifelse(debt_to_income > 0 & debt_to_income < 0.43, "within_limit", ifelse(debt_to_income == 0, "debt_free", "high")))


```

I think create a dataset with the variables I think could be valid for my model, and transform the relevant ones to factor variables.
```{r variable selection}
# create dataset with variables I want to use in my stepwise selection
conversions1 = conversions %>%
  dplyr::select(Approved, Gender, age_bracket, income_level, City_Category,
         Employer_Category1, Employer_Category2,
         Customer_Existing_Primary_Bank_Code, Primary_Bank_Type, Contacted,
         Source_Category, debt_level, Var1)

conversions1$Gender = as.factor(conversions1$Gender)
conversions1$age_bracket = as.factor(conversions1$age_bracket)
conversions1$income_level = as.factor(conversions1$income_level)
conversions1$City_Category = as.factor(conversions1$City_Category)
conversions1$Employer_Category1 = as.factor(conversions1$Employer_Category1)
conversions1$Employer_Category2 = as.factor(conversions1$Employer_Category2)
conversions1$Customer_Existing_Primary_Bank_Code =
  as.factor(conversions1$Customer_Existing_Primary_Bank_Code)
conversions1$Primary_Bank_Type = as.factor(conversions1$Primary_Bank_Type)
conversions1$Source_Category = as.factor(conversions1$Source_Category)
conversions1$Contacted = as.factor(conversions1$Contacted)
conversions1$debt_level = as.factor(conversions1$debt_level)


```

Next up, I use the stepwise variable selection process to identify statistically significant variables. Then I try model output as a logistic regression. 
```{r stepwise and logit}
# Stepwise selection to see which variables to use
null_state = lm(data=conversions1, Approved ~ 1)
full_state = lm(data=conversions1, Approved ~ .)

stepwise = stepAIC(null_state, scope=list(lower=null_state, upper=full_state),
                   direction="forward", na.rm=TRUE)

stepwise$anova
summary(stepwise)

# first model from the result of the stepwise selection
m1_logit = glm(Approved ~ income_level + Primary_Bank_Type + Employer_Category1 + 
    debt_level + Gender + age_bracket,
    data=conversions1, na.action=na.exclude)

summary(m1_logit)
anova(m1_logit, test="Chisq")

m1_logit

```


Now it's time to try the logit regression model on the test data. I first make the necessary transformations.
```{r logit test}
# converting DOB to age in years
test_df = test_df %>%
  mutate(DOB = as.Date(test_df$DOB, "%d/%m/%y"))

test_df$year <- format(as.Date(test_df$DOB), "%Y")
test_df$year = as.numeric(test_df$year)

date = test_df$DOB

test_df$age = 2018 - year(date)
test_df$age = abs(test_df$age)

test_df$age[is.na(test_df$age)] = 0

# binning age into 3 categories (based on "middle age" being considered 45-65)
test_df = test_df %>%
  mutate(age_bracket = ifelse(age >= 45 & age <= 65, "middle_aged",
                         ifelse(age >= 18 & age < 45, "young_adult", "senior")))
  

# binning income into 3 categories ($22,865.10 being the US federal govt for low income, and $100,000 being considered high income)
test_df = test_df %>%
  mutate(income_level = ifelse(Monthly_Income <= 1905, "low", 
                          ifelse(Monthly_Income > 1905 & Monthly_Income < 8334, "middle", "high")))

# Debt to income (after withdrawing current EMI)
test_df = test_df %>%
  mutate(income_post_EMI = Monthly_Income - Existing_EMI)

test_df = test_df %>%
  mutate(debt_to_income = test_df$income_post_EMI/test_df$Loan_Amount)

test_df$debt_to_income[is.na(test_df$debt_to_income)] = 0

# Binning of debt_to_income
test_df = test_df %>%
  mutate(debt_level = ifelse(debt_to_income > 0 & debt_to_income < 0.43, "within_limit", ifelse(debt_to_income == 0, "debt_free", "high")))

# converting to factor
test_df$Employer_Category2 = as.factor(test_df$Employer_Category2)
test_df$age_bracket = as.factor(test_df$age_bracket)
test_df$income_level = as.factor(test_df$income_level)
test_df$debt_level = as.factor(test_df$debt_level)


# PREDICT with LOGIT REGRESSION
test_df1 = test_df %>%
  mutate(pred_approved = predict(m1_logit, newdata=test_df, type='response'))
```

I had a decent ROC AUC with the logit model, but I wanted to see if this could be improved using a random forest model, so I decided to try that. I achieved a slightly better ROC AUC of 0.764.
```{r random forest}
sapply(conversions1, function(x) sum(is.na(x)))

m1_rf = randomForest(Approved ~ income_level + Primary_Bank_Type +
                       Employer_Category1 + debt_level + Gender + age_bracket,
                     data=conversions1)
m1_rf

test_df2 = test_df %>%
  mutate(approved = predict(m1_rf, newdata=test_df))
```

